---
title: "ggPicrust2"
output: html_document
date: "2025-05-06"
---

```{r, eval=FALSE}
rm(list = ls())
suppressMessages({
        library("tidyverse")
        library("phyloseq")
        library("rstatix")
        library("vegan")
        library("picante")
        library("data.table")
        library(dplyr)
        library(tidyr)
})

no_of_cores = 16
load("/work/benson/bpeng4/76Seeds_Peptides/data/intermediate/Peptide_ps.rda")
ls()
ps

###Pre-arrangement on  the phyloseq file
################
#Subset for only landrace and teosinte
ps <- subset_samples(ps, Category %in% c("Landrace", "Teosinte"))
ps.clean <- subset_taxa(ps, Kingdom == "Bacteria") %>%
  subset_taxa(!is.na(Phylum)) %>%
  subset_taxa(!Class %in% c("Chloroplast")) %>%
  subset_taxa(!Family %in% c("Mitochondria"))
ps.clean
#Filter Out Taxa Exist in At Least 25% samples
ps.clean.p0 <- filter_taxa(ps.clean, function (x) {sum(x > 0) >= 74}, prune=TRUE)
ps.clean.p0

#Average by Category
ps_melt <- psmelt(ps.clean.p0)  # Long-format data frame
# Average abundance for each taxon by Category.
group_var <- "Category" 
mean_abund_df <- ps_melt %>%
  group_by(OTU, !!sym(group_var)) %>%
  summarise(Abundance = mean(Abundance), .groups = "drop") %>%
  pivot_wider(names_from = !!sym(group_var), values_from = Abundance)
# Build New OTU table
otu_avg_mat <- as.matrix(mean_abund_df[,-1])
rownames(otu_avg_mat) <- mean_abund_df$OTU
otu_avg <- otu_table(otu_avg_mat, taxa_are_rows = TRUE)
#Subset tax_table and build sample_data
tax <- tax_table(ps)
tax <- tax[rownames(otu_avg_mat), ]
group_names <- colnames(otu_avg_mat)
sample_df <- data.frame(Group = group_names)
rownames(sample_df) <- group_names
sample_data_avg <- sample_data(sample_df)
#Assemble the new phyloseq file
ps_avg <- phyloseq(otu_avg, tax, sample_data_avg)
#Assign refseq to ps_avg
ps_avg <- merge_phyloseq(
  otu_table(ps_avg),
  tax_table(ps_avg),
  sample_data(ps_avg),
  refseq(ps.clean.p0)[taxa_names(ps_avg)]
)


```

```{r, eval=FALSE}
##Prepare Files for Picrust
#Difine Functions
phyloseq_counts_to_df <- function(ps) {
  df <- as.data.frame(t(otu_table(ps)))
  df <- cbind("#OTU ID" = rownames(df), df)  # Cheeky way to add a comment to first line
  
  # Note: we use OTU ID above since at the moment this function is meanly used
  # to pipe files into picrust, which needs the 'OTU ID'.
  return(df)
}

write_counts_to_file <- function(ps, filename) {
  df <- phyloseq_counts_to_df(ps)
  
  write.table(df,
              file=filename,
              sep = "\t",
              row.names = FALSE)
}

write_seqs_to_file <- function(ps, filename) {
  df <- as.data.frame(refseq(ps))
  
  unlink(filename)  # make sure we delete before we concatenate below
  
  names <- rownames(df)
  for (i in 1:nrow(df)) {
    cat(paste0(">", names[i]), file=filename, sep="\n", append=TRUE)
    cat(df[i,1], file=filename, sep="\n", append=TRUE)
  }
  message(paste0("Wrote ", nrow(df), " ASV sequences to file ", filename))
}

dir.create("picrust2", showWarnings = FALSE)
write_counts_to_file(ps.clean.p0, filename = "picrust2/raw_counts.tsv")
write_seqs_to_file(ps.clean.p0, filename = "picrust2/seqs.fna")
```


Average_Calculation <- function(ps.clean.p0) {
  # Ensure input is a phyloseq object
  if (!inherits(ps.clean.p0, "phyloseq")) {
    stop("Input must be a phyloseq object")
  }
  
  # Step 1: Merge samples by "Abbrv"
  ps_merged <- merge_samples(ps.clean.p0, group = "Abbrv")
  
  # Step 2: Get the number of original samples per Abbrv group
  sample_counts <- table(sample_data(ps.clean.p0)$Abbrv)
  
  # Step 3: Reorder sample_counts to match sample_names in ps_merged
  sample_counts <- sample_counts[sample_names(ps_merged)]
  
  # Step 4: Convert OTU table to matrix for manipulation
  otu_mat <- as(otu_table(ps_merged), "matrix")
  
  # Step 5: Average the counts by number of merged samples
  if (taxa_are_rows(ps_merged)) {
    otu_mat <- sweep(otu_mat, 2, sample_counts, "/")  # divide each column (sample)
  } else {
    otu_mat <- sweep(otu_mat, 1, sample_counts, "/")  # divide each row (sample)
  }
  
  # Step 6: Replace OTU table with the averaged one
  otu_table(ps_merged) <- otu_table(otu_mat, taxa_are_rows = taxa_are_rows(ps_merged))
  
  return(ps_merged)
}

Average_Calculation(ps.clean.p0)





```{bash, eval=FALSE}
module load picrust2/2.4
# Use an extra cd to go to your work directory (not shown)
cd picrust2

picrust2_pipeline.py -s seqs.fna -i raw_counts.tsv -p 2 -o results \
                     --stratified --verbose

cd /work/benson/bpeng4/76Seeds_Peptides/picrust2/results

add_descriptions.py -i EC_metagenome_out/pred_metagenome_unstrat.tsv.gz -m EC \
-o EC_metagenome_out/pred_metagenome_unstrat_descrip.tsv.gz

add_descriptions.py -i KO_metagenome_out/pred_metagenome_unstrat.tsv.gz -m KO \
-o KO_metagenome_out/pred_metagenome_unstrat_descrip.tsv.gz

add_descriptions.py -i pathways_out/path_abun_unstrat.tsv.gz -m METACYC \
-o pathways_out/path_abun_unstrat_descrip.tsv.gz

# Decompress tsv files
find . -name "*desc*" -exec pigz -dkv \{\} \;

cp ./picrust2/results/EC_metagenome_out/pred_metagenome_unstrat_descrip.tsv  ./picrust2/EC.tsv
cp ./picrust2/results/KO_metagenome_out/pred_metagenome_unstrat_descrip.tsv  ./picrust2/KO.tsv
cp ./picrust2/results/pathways_out/path_abun_unstrat_descrip.tsv  ./picrust2/path.tsv
```

#Enzyme Commission number(EC) Output
```{r, eval=FALSE}
setwd("/work/benson/bpeng4/76Seeds_Peptides")
p2_EC = paste0("picrust2", "/EC.tsv")
p2EC = as.data.frame(fread(p2_EC))
rownames(p2EC) = p2EC$"function"
p2EC = p2EC %>%
        mutate(across(-c(`function`, description), round ))
p2EC = as.matrix(p2EC[,-1]) # drop off function column as we have it as row names
```

#KEGG Orthology(KO) Output
```{r, eval=FALSE}
p2_KO = "/Users/bopeng/Documents/76Seeds_Peptides/picrust2/KO.tsv"
p2KO = as.data.frame(fread(p2_KO))
rownames(p2KO) = p2KO$"function"
p2KO = p2KO %>%
        mutate(across(-c(`function`, description), round ))
p2KO = as.matrix(p2KO[,-1])
```

####ggpicrust2
#library packages
```{r}
library(magrittr)
library(ggpicrust2) 
library(tibble) 
library(fgsea)
library("enrichplot")
library("ComplexHeatmap")
library("circlize")
```

#Visualization for Gene Set Enrichment Analysis(GSEA) results
```{r}
# Prepare abundance data
abundance_data0 <- as.data.frame(p2KO) 
abundance_data0 <- abundance_data0[, -1]

# Prepare metadata
metadata0 <- ps.clean.p0@sam_data

#Change values to numeric values in the dataframe
rn <- rownames(abundance_data0)
cn <- colnames(abundance_data0)
abundance_data0 <- as.data.frame(lapply(abundance_data0, function(x) as.numeric(as.character(x))))
rownames(abundance_data0) <- rn
colnames(abundance_data0) <- cn


# Run GSEA analysis 
gsea_results <- pathway_gsea( abundance = abundance_data0, 
                              metadata = metadata0,
                              group = "Category", 
                              pathway_type = "KEGG", 
                              method = "fgsea"
                              )

## Annotate results（Error : object "KEGG" not found
annotated_results <- gsea_pathway_annotation( 
    gsea_results = gsea_results,
    pathway_type = "KEGG"
    )
                              
# Create enrichment plot
visualize_gsea(gsea_results, plot_type = "enrichment_plot", n_pathways = 10)

# Create dotplot
visualize_gsea(gsea_results, plot_type = "dotplot", n_pathways = 20)

# Create barplot
visualize_gsea(gsea_results, plot_type = "barplot", n_pathways = 15)

# Create network plot
visualize_gsea(gsea_results, plot_type = "network", n_pathways = 15)

# Create heatmap 
visualize_gsea(gsea_results,
                plot_type = "heatmap",
                n_pathways = 15,
                abundance = abundance_data0, 
                metadata = metadata,
                group = "Environment"
                )

```

